Modal Verification Patterns for Systems

### 2025-06-23 paper.tex
+

Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}. Like VMMs 
These snapshots overlap on the unmodified portions of the filesystem (the left-most grid section), with the reclaimed the chunks (the middle gray section) becoming available for reuse once the new snapshot becomes the official current state of the filesystem. This permits the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk. Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree. Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$. This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs. 
An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk. Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures. 
\begin{figure}
    \includegraphics[width=\columnwidth]{journalling.png}
    \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
    \label{fig:journal}
\end{figure}

\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}
\todo[inline,color=cyan]{This looks like it's verbatim copy-pasted from the ModFS draft; if this exact text
it used in this publication, it means we'd need to completely rewrite the relevant text in that draft.}
With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}. Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle. Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification as abstarcted in Figure \ref{fig:journal}) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location. This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}

>>>>

Like VMMs navigating between different address-spaces, a \textsf{CoW} filesystem switches from either the old to the new root yields consistent views of the filesystem at different points in time, each commonly referred to as a \emph{snapshot}. Like VMMs loading a process's memory mappings to be the current view of the memory by setting the control register to show the unique root address of the page-table tree of the process, the filesystem to atomically switch between consistent on-disk states, by updating the oldest so-called \emph{super} chunk at a fixed location on disk. Super chunks, as we see in Figure \ref{fig:cowindex} as ($\textsf{Super}_0$ and $\textsf{Super}_1$), are the first two blocks of the disk dedicated to keep the state of the file system which is composed of some meta-data on the filesystem, and most importantly to the root chunk of the file system index-tree. 
\begin{figure}
    \includegraphics[width=\columnwidth]{journalling.png}
    \caption{A Filesystem Utilizing \emph{Journalling} for Consistent Disk-Layout ~\cite{chajed2021gojournal}}
    \label{fig:journal}
\end{figure}

\paragraph{Different Models in Ensuring Disk Consistency of Filesystems}
\todo[inline,color=cyan]{This looks like it's verbatim copy-pasted from the ModFS draft; if this exact text
  it used in this publication, it means we'd need to completely rewrite the relevant text in that draft.}
Therefore, in case of a crash during the update to the \textsf{file "foo"}, we can reboot the file system from the previous persisted root, \textsf{root chunk}, via reading the $\textsf{Super}_0$. This means the whole disk consistency checks are unnecessary when recovering from a crash~\cite{bonwick2003zettabyte,hitz1994file}; recovery is immediate, as opposed to requiring hours or days~\cite{henson2006reducing,agrawal2009speedy} (or longer~\cite{henson2006chunkfs}) for recovering a large volume’s consistency from logs. 
An important additional subtlety with this kind of filesystem is \emph{when} blocks in the reclaimed region can actually be reused: only \emph{after} the snapshot that made those blocks ``free'' has been completely written to disk. Allowing reuse before this could result in in-place updates: part of the new snapshot could attempt to store data in space used by the prior snapshot, allowing inconsistent on-disk structures.

With their intricate semantics and critical place in modern systems, formal verification of functional correctness for storage systems (including filesystems and other kinds of storage) have drawn much attention~\cite{ntzik-gardner-15,ntzik2018concurrent,chen-crash-16}\todo{lots more cites}. Unlike our \textsc{CoW} filesystem \textsf{ModFs} which allocates new chunks and rewrites all the nodes from leaf to root on the write-amplification path, some filesystems allows in-place updates to handle the heavy workloads more efficiently in principle. Unlike \textsc{CoW} filesystems, filesystems with in-place updates need to implement some form of \emph{journalling} to ensure that a crash partway through an in-place update can be rolled back (using an undo journal that records the original contents before modification as abstarcted in Figure \ref{fig:journal}) or completed (using a write-ahead journal that first writes new data in unused space before \emph{also} updating in place).\footnote{\textsf{CoW} filesystems can be viewed as a twist on write-ahead journals that uses the journal as the actual data once written to a fresh location, thereby reclaiming the original location. This results in a performance and complexity trade-off: journalling writes twice for every updated block and requires recovery logic to roll back or re-execute the log, while \textsf{CoW} techniques write each update only once and require slightly more intricate allocation strategies to compact fragmentation.}
